---
title: "Initial Data Exploration"
output: 
  html_notebook: default
  word_document: default
---

## Purpose

This document does a very cursory exploration of the data included for the health care (HC) data files.  There are three and each will have its own section.

# HC Chronic

```{r load_packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
```
```{r load_chronic, echo=FALSE, message=FALSE, warning=FALSE}
HC_Chronic <- read_excel("C:/Users/l2eve/OneDrive/Documents/repo/dat_410/data/health_care_files_for_final_project/HC_Chronic.xlsx")
```
```{r show_truncated_table, echo=FALSE, message=FALSE, warning=FALSE}
short_chronic <- head(HC_Chronic) %>%
  select(-`2012`, -`2011`, -`2010`, -`2009`, -`2008`, -`2007`) %>%
  knitr::kable()
```
Below is a truncated table to help visualize the raw data that is provided.

`r short_chronic`

The columns included are:
```{r list_columns, echo=FALSE, results='asis'}
chronic_col <- colnames(HC_Chronic)
cat(paste('-', chronic_col), sep = '\n')
```

The columns with year indicators likely have the calculation '*characteristic multiplier factors*' which is the value we will eventually be forecasting.  The ID column is a unique identifier for each row and won't really be of use during the analysis.  Data set is an indicator for which data set it comes from Chronic, Insurance, or Vaccinations.  This will only become useful should the analysis require all three data sets to be appended.  This leaves Label, Characteristic, and Chronic Condidtions as the dimensions that will drive the analytic.

To that end, each of these dimension contains the following values:

```{r label_values, echo=FALSE, results='asis'}
HC_Chronic %>%
  select(Label) %>%
  unique() %>%
  knitr::kable()
```

```{r characterisic_values, echo=FALSE, results='asis'}
HC_Chronic %>%
  select(Characteristic) %>%
  unique() %>%
  knitr::kable()
```

```{r chronic_value, echo=FALSE, results='asis'}
HC_Chronic %>%
  select(`Chronic Conditions`) %>%
  unique() %>%
  knitr::kable()
```

The case study says of the chronic data:

For chronic illness, provide the top three label/characteristics for each group: 1 chronic illness,
2–3 chronic illness, and 4+ chronic illness. For the identified top three, generate a multiplier and
coefficient by forecasting the next value using 2002 through 2012.

This means to me, we'll need to decide on how to quantify and rank these groups using the yearly multiplier and coefficient.  Some initial ways that this might be done are to just sum the coefficients for the entire time frame, use an average, use a mean, use a median, or even use a weighted mean.

```{r data_munging, echo=FALSE, message=FALSE, warning=FALSE}
chronic.long <- HC_Chronic %>%
  pivot_longer(c(-ID, -`Dataset`, -Label, -`Chronic Conditions`, -Characteristic), names_to = "year")

chronic.summary <- chronic.long %>%
  group_by(Label, Characteristic, `Chronic Conditions`) %>%
  summarize(summed = sum(value, na.rm = TRUE),
            avg = mean(value, na.rm = TRUE),
            med = median(value, na.rm = TRUE),
            n.val = n()) %>%
  arrange(`Chronic Conditions`, Label, Characteristic)
```
```{r chronic.boxplot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(chronic.long, aes(x = `Chronic Conditions`, y = value, fill = Label)) +
  geom_boxplot(outlier.color = "red", outlier.alpha = 0.5) +
  ggthemes::theme_tufte() +
  ggthemes::scale_fill_hc()
```
```{r view_sum_tbl, echo=FALSE, message=FALSE, warning=FALSE}
knitr::kable(chronic.summary)
```

